lfl: An \R{} Package for Linguistic Fuzzy Logic

### 2021-06-09 article.tex
Paper update pre-case study

The aim of this paper is to present a new package for the \R{} statistical environment \citep{R2020, Rintro} that enables the use of the linguistic fuzzy logic in data processing applications. The package provides implemented tools for using the results of original work of \cite{Novak08, Novak:PbLD, DvoStep:PbLD2015, StepBurda:FRBE_FSS}, and others, and it provides executable routines that are not freely available anywhere else.

@@

The \pkg{lfl} package also provides functions for searching for \emph{fuzzy association rules} \citep{srikant95}. Together
with PbLD, they can be used as a machine learning tool for classification or regression problems, see \cite{KupkaRusnok:Regression_ESWA_INS17}.
The package also includes the \emph{Fuzzy Rule-based Ensemble}, a tool for time series forecasting
\citep{StepBurda:FRBE_FSS}, which is built on top of the fuzzy association rules search algorithm and PbLD.

@@

@@

@@

>>>>

The aim of this paper is to present particular package for the \R{} statistical environment \citep{R2020, Rintro} named {\bf lfl} that enables the use of the linguistic fuzzy logic in data processing applications. The package provides implemented tools for using the results of original work of \cite{Novak08, Novak:PbLD, DvoStep:PbLD2015, StepBurda:FRBE_FSS}, and others, and it provides executable routines that are not freely available anywhere else.

@@

The \pkg{lfl} package also provides functions for searching for \emph{fuzzy association rules} \citep{srikant95}. Together
with PbLD, they can be used as a machine learning tool for classification or regression problems, see \cite{KupkaRusnok:Regression_ESWA_INS17}.
The package also includes the \emph{Fuzzy Rule-based Ensemble} (FRBE), a tool for time series forecasting
\citep{StepBurda:FRBE_FSS}, which is built on top of the fuzzy association rules search algorithm and PbLD.

@@

\Martin{checkni prosím tento odstavec po faktografické stránce}

It is important to note that {\bf lfl} package was firstly issued in {\color{red}2014} and described in \cite{Burda_lfl:FUZZ15}. The differences are, however, essential. Indeed, many of the above-mentioned results that are implemented in the current version of {\bf lfl} were not even published and thus, could hardly be incorporated in the original version. In particular, e.g., algebras for partial fuzzy logics including the novel structures such as Dragonfly algebras or lower estimation algebras; fuzzy relational calculus including the extensions such as the use of generalized quantifiers, excluding features, or unavoidable features. Furthermore, apart from these latest results, it newly includes even fundamental foundations such as the basic  fuzzy relational compositions that allow to employ the extensions and also to deal with fuzzy inputs when incorporating fuzzy rule-based systems, and regarding the latter notion, the standard fuzzy relational models (Mamdani-Assilian and the implicative one) have been added as well including the related defuzzification techniques. This changes the original nature of {\bf lfl} package from an R-package implementation of the linguistic control software LFLC, see \cite{dvo:lflc} with a few extensions (associations rules and FRBE), to a brand new complex package that can be used for distinct purposes and for building more complex tools due to the presence of sound mathematical foundations of fuzzy modeling techniques.

@@

The \pkg{lfl} package is distributed under the terms of the GNU General Public License (GPL), which guarantees the user the freedom to use it, study, share and modify.

@@

\section{Applicability}
\label{sec:applicability}

This section places the \pkg{lfl} package into the perspective of the others from the practical applicability point of view. Firstly, we briefly recall the other tools and packages, secondly, we present a real case-study problem.

\subsection{Other related packages}

As mentioned above, \pkg{lfl} is by far not the only SW or even R-package related to the fuzzy methods. An exhaustive study on SW related to fuzzy techniques can be found in \cite{AlcalAlonso:fuzzySW_IEEETFS16}. Staying in the environment of R-packages, the above-mentioned \pkg{frbs} \citep{frbsPkg,FRBS:jstatsoft14} is, according to the subjective opinion of the authors, topically the closest one and it seems to be also developed to the widest applicable potential.

Other packages that deal with fuzzy rules are, for instance, \pkg{FisPro} \citep{fisproPkg}, \pkg{FuzzyToolkitUoN} \citep{FuzzyToolkitUoNPkg} or \pkg{FuzzyR} \citep{FuzzyRPkg} where the first two ones focuse on Mamdani-Assilian and Takagi-Sugeno systems which makes them closer to \pkg{frbs} while the latter one focuses mainly on the well-known neurofuzzy system ANFIS.

The portfolio of problems possibly solvable by existing R-packages with fuzzy techniques is, indeed, much wider. We may find, for example, packages for fuzzy linear regression, AHP, fuzzy statistics, or fuzzy decision trees and random forests \citep{fuzzyforestPkg,fuzzyforestJSS}.

If we intend to compare the \pkg{lfl} package with the most related ones from the area/technique coverage point of view, we find the following differences and similarities.

Due to the specific character of the PbLd inference and the related necessity of the use of evaluative linguistic expressions \citep{Novak:ELE_FSS} based on the basic trichotomy, none of the existing packages contains these structures. Neither the predefined fuzzy sets modeling expressions created by the application of linguistic hedges to the basic triplet, nor the related implication-based inference based on a sort of ``pre-selection'' of rules. This fact that is not that surprising, however, up to the best knowledge of the authors, none of the existing packages contains fuzzy relational model of implicative rules. Having in mind its importance, this contribution of {\bf lfl} to the existing tools is essential.

For the sake of the completeness of the \pkg{lfl} package and for the sake of the consistency of the implementation with the implicative rules, \pkg{lfl} contains also Mamdani-Assilian model of fuzzy rules. However, here we have to note that this approach is already employed in several packages, e.g., in \pkg{FisPro}, \pkg{FuzzyToolkitUoN}, and mainly in \pkg{frbs}. We are also emphasizing  distinct efficient and effective modifications and learning techniques, e.g., neural-network based fuzzy inference system ANFIS employed in \pkg{FuzzyR} and in \pkg{frbs}. In the latter package, ANFIS is  supplemented by DENFIS, hybrid neural approach HYFIS, or distinct genetic (evolutionary) approaches, e.g. by genetic fuzzy systems based on Thrift and MOGUL methods. This, jointly with other inferences (Takagi-Sugeno-Kang or Ishibuchi's method) make \pkg{frbs} a powerfull and a rich package for distinct purposes, such as regression, data-driven classification, control, or decision-making.

Owing to mention other differences, \pkg{lfl} contains complete and very recent group of fuzzy relational composition calculus including the most recent techniques \citep{CaoStep:ExcludfeatureESWA,Step_etal:Unavoidable_KNOSYS2020}. These structures can be used in the inference processing fuzzy inputs as well as in distinct, e.g., expert-driven, classification or decision-making tasks. We also emphasize that the implementation allows to compose partial fuzzy relations \citep{Step_etal_Dragon_IJAR2019}.

Generally, the incorporation of algebraic structures is different in \pkg{lfl}, it contains the main three residuated structures that are used as parameters \Martin{je to tak, uvádí se to tak anglicky, že jsou ty agebry "parametry"?}. Furthermore, for the sake of the ability to deal with partial fuzzy logics, it contains six partial algebraic structures that can be combined with any of the three basic underlying residuated structures.

The \pkg{lfl} package also contains fuzzy association rules mining algorithm that generates linguistic rules to which the PbLD inference is tailored. The mining algorithm works with genuine approach that reflects membership degrees and that was used for the generation of the fuzzy rule base ensemble technique for time series forecasting \citep{StepBurda:FRBE_FSS}. This ensemble is also employed in \pkg{lfl}.

Association rules have rich history, before they got their name in \citep{Agrawal:assoc_rules}, they were propoed as so-called GUHA method already in \citep{Hajek:GUHA}. However, we are not aware of fuzzy associations rule mining algorithm implemented in any existing R-package that would be any reflecting membership degrees. We may recall, e.g., \pkg{arules} package, however, it deals with the crisp variant of associations rule mining only.

The survey provided above brings a comparison of the R-packages that are topic-closest to \pkg{lfl} however, its goal is not to compare them from the superiority point of view. This is probably even impossible. The goal is to compare them from the area and technique coverage point of view, to show the similarities and differences. The fact is that the packages have rather complementary positions not duplicating each other. Out of the existing packages, we would like to once more emphasize the \pkg{frbs} package that provides the richest set of methods for distinct purposes (control, regression, classification) -- many of them not even attempted in \pkg{lfl}. The \pkg{lfl} package is closer to more mathematically oriented foundations in providing algebraic backgrounds and tools from the fuzzy relational calculus that may be useful for distinct purposes. In the area of  inference systems, it complements existing packages with fundamental implicative rule bases and specific PbLD method, that is even followed by associations mining and FRBE for time series forecasts.

\subsection{Case study I -- expert classification of amphibians}

This section briefly describes the application of fuzzy relational compositions to the expert classification of animal species, namely of amphibians. It nicely demonstrates the power of chaining distinct composition blocks to their extensions. We avoid the details and the experimental comparison with classical data-driven classifiers as this can be found int he very recent source \cite{Step_etal:Unavoidable_KNOSYS2020} and we concentrate on the use of the R-package.

The problem setting is as follows. We are given 79427 training data samples (animal records) which constitutes the set $X$. Each data sample is a vector containing 29 values -- the features constituting the set $Y$. These are, e.g., Boolean descriptors or  distinct colors from the top or from the bottom of an animal. The task is to classify each animal into one of the 21 species -- which constitute the set of classes $Z$.

Expert knowledge is then encoded in the matrix $S$ declaring the ``supportive'' features for given species; in the matrix $E$ declaring the excluding features for particular classes; and in matrix $U$ declaring the unavoidable features for given classes. All these matrices (mathematically fuzzy relations on $Y\times Z$) are naturally of the dimension $29\times 23$ and were determined by a biologist. Note, that for the testing reasons, other 79418 data samples were used in \cite{Step_etal:Unavoidable_KNOSYS2020}.

The results of chosen compositions for a fixed row from matrix $R$, i.e., for a particular animal record, are vectors of 23 values -- each value expressing a membership degree to a particular species. However, it is important to note that it is not a membership degree of the given animal to the given species as each animal belongs to a single species only.

For example, if we use the command:
\begin{Sinput}
a <- algebra("lukasiewicz")
compose(R, S, alg = a, type = "basic")
\end{Sinput}
we get $(R\circ S)(x,z)$ which is the truth degree of the statement \emph{``there exists at least a single feature that is carried by animal record $x$ and it belongs to the features related supporting the suspicion for species $z$''}. As we can see from \cite{Step_etal:Unavoidable_KNOSYS2020}, this composition would lead to 100\% accuracy (sensitivity) in a sense of true positive cases however, very lows accuracy (specificity) in a sense of false positive cases. Indeed, on average over the whole testing set, there were 18 vector values equal to 1 and so, 18 species were found suspicious on average.

Right for the reasons of increasing the true negative accuracy and thus, narrowing the set of suspicious species, the other extensions may be used. For example, adding the commands:
\begin{Sinput}
RE <- compose(R, E, alg = a, type = "basic")
a$pt(RS, a$n(RE))
\end{Sinput}
leads to $(R\circ S^{\backprime}E )(x,z)$ which is the truth degree of the statement \emph{``there exists at least a single feature that is carried by animal record $x$ and it belongs to the features related supporting the suspicion for species $z$ and at the same time, $x$ carries no feature that would be excluding for the species $z$''}. Such an extension directly reduces the number of suspicious species to 3.41 on average while it keeps the 100\% accuracy in the true positive cases point of view, see \cite{Step_etal:Unavoidable_KNOSYS2020}.

Furthermore, we can involve also the unavoidable features:
\begin{Sinput}
RU <- compose(R, U, alg = a, type = "super")
a$pt(RS, a$n(RE), RU)
\end{Sinput}

bla bla bla

\subsection{Case study II -- fuzzy associations rules from the Iris data}

\Martin{zatim nemam co psat, dokud nebudou obrysy experimentu}

The second case study will work with the famous Iris dataset \citep{iris1, iris2}:

### 2021-06-09 cover.tex
Paper update pre-case study

@@

\begin{document}
Cover letter

@@

\todo{answer}

@@

\todo{answer}

@@

\todo{answer}

@@

\todo{answer}

@@

\todo{answer?}

@@

\todo{answer}

@@

{ \bf Thank you, we have added reference to the suggested web page into the paper.}

{ \it Language:
\begin{itemize}
    \item p.1: Up to the best knowledge -- To the best of the authors' knowledge; there is NO other
    \item p.2: was initially developed IN [2]
    \item p.3: PbLd -- pBLD
    \item p.5: forms A residuated lattice
    \item p.10: Jan ŁUKASIEWICZ
    \item p.11: simNply
    \item p.18: let cough EXCLUDE
    \item p.22: "crucial thing" is too informal
    \item p.34: breaks-points -- break-points
    \item p.45: in such A case; the perception ALSO fires THE rules
\end{itemize} }

{ \bf Thank you for pointing out that typos -- we have fixed them as suggested. }

@@

\todo{answer}

@@

\todo{answer}

>>>>

\newcommand{\separator}{\rule{\textwidth}{0.4pt}}

@@

\begin{document}

\section*{FSS-D-21-00085 (R1) - COVER LETTER}

Dear editors, dear reviewers,

we are grateful for your remarks and comments including the critical part of the assessment of our work FSS-D-21-00085 that helped us to improve the article. In the subsequent text, we address your concerns point by point.

@@

{RE: Thank you. Indeed, the preliminary version published in FUZZ-IEEE 2015 could not contain all. Not only that the page extent of the current submission is maybe three times bigger, but mainly the package itself evolved significantly. The original version did not contain all algebras, it had no algebras for partial fuzzy logics, it had no compositions of fuzzy relations. Consequently, the extensions (excluding features, unavoidability, generalized quantifiers) were not there either -- neither in the package, nor in the paper from 2015. Naturally, these parts were not even existing from the theoretical point of view in 2015 so, the novelty is indeed present. Note that just these extension are bigger than what is the usual ``size'' of many other (fuzzy related) R-packages. Moreover, the original conference paper from 2015 indeed did not have an extent to sufficiently describe even those functions that were already existing so, we are convinced that this article is at least 80\% different and cannot be identified with the original one.

But we comprehend and we have to agree with the reviewer, that the differences should be mentioned in the paper. It is our duty, not the duty of the reviewer nor of the reader. We accept this remark, we apologize, and we modify the text accordingly. It is reflected mainly in the last part of Introduction, right before Section 1.1 starts.}
\separator

@@

{RE: Thank you. Even here we have to admit we were ourselves constructing several variants on how the paper should look like and we had a feeling that a similar opinion may occur. On the other hand, the Reviewer 2 stated ``A well-written paper introducing the lfl (linguistic fuzzy logic) package for R, which is amongst the most popular environments for scientific computing $\ldots$ The manuscript is of appropriate length — it describes the package features very comprehensively, provides commands to be run by a practitioner, gives details on the mathematics behind the algorithms. R commands and their outputs are an integral part of the paper and demonstrate the typical package use cases. They can be considered as something more than numbered examples/illustrations in more mathematical works, as they can be read as text paragraphs too. They also explain the API design.
Also they illustrate the usefulness of the package in statistical analysis activities. {\bf This is definitely a research paper.}'' And this was our second feeling we had. These two feelings were fighting in our heads so, we comprehend opinions of both reviewers.

Therefore, and in order to find the best form of the paper, we have contacted the Editors-in-Chief, with a question regarding the form of such a paper fitting the scope of the journal. We all agreed that pure SW manual is not the desirable form. We have been advised to get motivated by the article {\it M. B. Ferraro, P. Giordani, A toolbox for fuzzy clustering using the R programming language, Fuzzy Sets and Systems 279 (2015) 1–16} that carries the desirable form. We are convinced we kept a very similar style in order to submit a paper consistent with the journal's scope. Your comments below convinces us we can do more, so, we have mirrored it in the revised version. Thanks a lot.}

\separator

@@

{RE: Thanks a lot, as stated above, we respect this proposal. The comparison is mirrored in the text constituting Section 7.1.}

\separator

@@

{RE: Thanks a lot, as stated above, we respect this proposal. Following your suggestion, we have added Section 7.2 containing the proposed case study. {\color{red}Toto ovšem musíme dodělat, článek visí vlastně jen na tomto}}

@@

{RE: Thank you very much for your judgement and your suggestion. Some of them are probably much more general, beyond the goals of this article, but we take them into account.}

\separator

@@

{RE: Thank you very much. Some of the above-mentioned packages appeared in the newly added comparison so, comments and suggestions of Reviewer 1 as well as Reviewer 2 were employed in a compact form.}

\separator

@@

RE: Thank you, we have added reference to the suggested web page into the paper.

\separator

{ \it Language:
\begin{itemize}
    \item p.1: Up to the best knowledge -- To the best of the authors' knowledge; there is NO other
    \item p.2: was initially developed IN [2]
    \item p.3: PbLd -- pBLD
    \item p.5: forms A residuated lattice
    \item p.10: Jan ŁUKASIEWICZ
    \item p.11: simNply
    \item p.18: let cough EXCLUDEG
    \item p.22: "crucial thing" is too informal
    \item p.34: breaks-points -- break-points
    \item p.45: in such A case; the perception ALSO fires THE rules
\end{itemize} }

RE: Thank you for pointing out that typos, we have corrected them.

\separator

@@

RE: Thank you for pointing this out -- we have fixed the code as suggested.

\separator

@@

The lfl package is distributed under the terms of the GNU GPL license. We have added to the paper a note about it.
